PCRB

### 2013-04-11 doc/PCRB.tex
Latest intermediate

The model takes the form
\begin{subequations}
	\label{eqn:model_general}
\begin{align}
	\vec x_k &= \vec f_k(\vec x_{k-1}) + \vec v_k \\
	\vec y_k &= \mat H\vec x_{k-1} + \vec w_k
\end{align}
\end{subequations}
where $\vec v_k$ and $\vec w_k$ are white, uncorrelated, zero mean Gaussian noise vectors with covariances of $\mat Q$ and $\mat R$, respectively. For this model the transition density is defined as
\begin{align}
	p(\vec x_k|\vec x_{k-1}) &= \mathcal N(\vec f(\vec x_{k-1})),\mat Q) \\
	&= (2\pi)^{-d/2} |\mat Q|^{-1/2} \exp\left(-\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \right)
\end{align}
and the likelihood is
\begin{align}
	p(\vec y_k|\vec x_k) &= \mathcal N(\mat H\vec x_k,\mat R) \\
	&= (2\pi)^{-n/2} |\mat R|^{-1/2} \exp\left(-\frac{1}{2}(\vec y_k - \mat H\vec x_k)^T \mat R^{-1} (\vec y_k - \mat H\vec x_k) \right)
\end{align}

@@

The posterior Cram\'er-Rao bound (PCRB) is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
\begin{align}
	\mat J &= \mathbb E\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right]
\end{align}
where the expectation is taken over $\vec y$ and $\vec x$. The PCRB bounds the covariance of an unbiased estimator, $\mat P$,
\begin{align}
	\mat P &\ge \mat J^{-1}
\end{align}

@@

An alternative formulation is to recursively compute the bound instead of the matrix, i.e., $\mat P_k = \mat J_k^{-1}$ directly rather than $\mat J_k$. The formulas suggested in \cite{Bergman2001} are applicable to wider class of models (e.g. multiplicative noise), denoted by a general transition function $\tilde{\vec f}(\vec x_k,\vec v_k)$. The recursion is
\begin{subequations}
\begin{align}
	\mat P_{k|k-1} &= \bar{\mat F}_k\mat P_{k-1|k-1}\bar{\mat F}_k + \bar{\mat G}_k\mat Q_k \bar{\mat G}_k^{-1} \\
		\mat P_{k|k} &= \left(\mat P_{k|k-1}^{-1} + \bar{\mat R}_{k-1}^{-1}\right)^{-1}
\end{align}
\end{subequations}
where
\begin{align}
	\bar{\mat F}_k &= \mathbb E \left[ \nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
	\bar{\mat R}_k^{-1} &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) \right] \\
	\bar{\mat G}_k &= \mathbb E \left[ \nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
	\mat Q_k^{-1} &= -\mathbb E \left[ \nabla_{\vec v_k}\nabla_{\vec v_k}^T \log p(\vec v_k) \right]
\end{align}
and $\mat P_0^{-1}$ is the inverse of the prior covariance. The model in \eqref{eqn:model_general} is represented by $\tilde{\vec f}(\vec x_k,\vec v_k) = \vec f(\vec x_k) + \vec v_k$. This simplifies the derivatives to
\begin{align}
	\nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat F_k(\vec x_k) \\
	\nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat I_{d\times d} \\
	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= -\mat H^T\mat R^{-1} \mat H 
\end{align}
and the corresponding matrices to	
\begin{align}
	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \label{eqn:pcrb_term_F}\\
	\bar{\mat R}_k^{-1} &= \mat H^T\mat R^{-1} \mat H \\
	\bar{\mat G}_k &= \mat I_{d\times d} 
\end{align}
Notice that $\mat Q^{-1}$ is not directly required in the recursion, so this method is applicable to neural field models where process noise is only added to a single state.

@@

@@

Let $\vec x_k = [a_k\quad b_k]^T$ where $a_k$ is the pendulum angle and $b_k$ is the angular velocity,
\begin{align}
	\vec f(\vec x_{k}) &= \left[\begin{array}{c}
								a_k + b_k\Delta \\ -b_k - g\sin(a_k)\Delta
								\end{array}\right]
\end{align}
The linearisation used for the EKF is
\begin{align}
	\mat F(\vec x_{k}) &= \left[\begin{array}{cc}
					1 & \Delta \\
					-g\cos(a_k) & 1 \\
				\end{array}\right]
\end{align}
The observation matrix is $H=[1\quad 0]$, so that only the first state element is measured.

>>>>

The model takes the form
\begin{subequations}
	\label{eqn:model_general}
\begin{align}
	\vec x_k &= \vec f_{k-1}(\vec x_{k-1}) + \vec v_k \\
	\vec y_k &= \mat H\vec x_k + \vec w_k
\end{align}
\end{subequations}
where $\vec v_k$ and $\vec w_k$ are white, uncorrelated, zero mean Gaussian noise vectors with covariances of $\mat Q$ and $\mat R$, respectively. For this model the transition density is defined as
\begin{align}
	p(\vec x_k|\vec x_{k-1}) &= \mathcal N(\vec f(\vec x_{k-1})),\mat Q) \\
	&= (2\pi)^{-d/2} |\mat Q|^{-1/2} \exp\left(-\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \right)
\end{align}
and the likelihood is
\begin{align}
	p(\vec y_k|\vec x_k) &= \mathcal N(\mat H\vec x_k,\mat R) \\
	&= (2\pi)^{-n/2} |\mat R|^{-1/2} \exp\left(-\frac{1}{2}(\vec y_k - \mat H\vec x_k)^T \mat R^{-1} (\vec y_k - \mat H\vec x_k) \right)
\end{align}

\section{Classical CRB}

The Fisher information matrix is
\begin{align}
	\mat J_{\rm F}(\vec x) &= \mathbb E_{\vec y| \vec x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec x) \right]
\end{align}
Notice only the likelihood is used since no prior statistics exist for the unknown parameters/state $\vec x$.

@@

The posterior Cram\'er-Rao bound (PCRB) is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
\begin{align}
	\mat J &= \mathbb E_{x,y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right] \\
	&= \mathbb E_{x,y}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec x) \right] + \mathbb E_{x}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec x) \right]
\end{align}
where the expectation is taken over $\vec y$ and $\vec x$ in the first term and $\vec x$ for the second term. The PCRB bounds the covariance of an unbiased estimator, $\mat P$,
\begin{align}
	\mat P &\ge \mat J^{-1}
\end{align}

@@

An alternative formulation is to recursively compute the bound instead of the matrix, i.e., $\mat P_k = \mat J_k^{-1}$ directly rather than $\mat J_k$. The formulas suggested in \cite{Bergman2001} are applicable to wider class of models (e.g. multiplicative noise), denoted by a general transition function $\tilde{\vec f}(\vec x_k,\vec v_k)$. The recursion is
\begin{subequations}
\begin{align}
	\mat P_{k|k-1} &= \bar{\mat F}_k\mat P_{k-1|k-1}\bar{\mat F}_k + \bar{\mat G}_k\mat Q_k \bar{\mat G}_k^{-1} \\
		\mat P_{k|k} &= \left(\mat P_{k|k-1}^{-1} + \bar{\mat R}_{k-1}^{-1}\right)^{-1}
\end{align}
\end{subequations}
where
\begin{align}
	\bar{\mat F}_k &= \mathbb E \left[ \nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
	\bar{\mat R}_k^{-1} &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) \right] \\
	\bar{\mat G}_k &= \mathbb E \left[ \nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
	\mat Q_k^{-1} &= -\mathbb E \left[ \nabla_{\vec v_k}\nabla_{\vec v_k}^T \log p(\vec v_k) \right]
\end{align}
and $\mat P_0^{-1}$ is the inverse of the prior covariance. The model in \eqref{eqn:model_general} is represented by $\tilde{\vec f}(\vec x_k,\vec v_k) = \vec f(\vec x_k) + \vec v_k$. This simplifies the derivatives to
\begin{align}
	\nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat F_k(\vec x_k) \\
	\nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat I_{d\times d} \\
	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= -\mat H^T\mat R^{-1} \mat H 
\end{align}
and the corresponding matrices to	
\begin{align}
	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \label{eqn:pcrb_term_F}\\
	\bar{\mat R}_k^{-1} &= \mat H^T\mat R^{-1} \mat H \\
	\bar{\mat G}_k &= \mat I_{d\times d} 
\end{align}
Therefore the final recursion is
\begin{align}
	\mat J_{k|k} = \left( \mat Q + \bar{\mat F}_{k-1} \mat J_{k-1}^{-1} \bar{\mat F}_{k-1}^T\right)^{-1} + \mat H_k^T \mat R^{-1} \mat H_k
\end{align}

Notice that $\mat Q^{-1}$ is not directly required in the recursion, so this method is applicable to neural field models where process noise is only added to a single state.

@@

\section{Hybrid CRB}

In this formulation the unknown vector, $\vec x$ is composed of random unknown elements, $\vec z$, and deterministic unknown elements, $\vec \theta$. The hybrid Fisher information matrix is
\begin{align}
	\mat J &= \mathbb E_{y,z|\theta}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec z|\vec\theta) \right] \\
	&= \mathbb E_{y,z|\theta}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec z,\vec\theta) \right] + \mathbb E_{z|\theta}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec z|\vec\theta) \right] \\
	&= \mathbb E_{z|\theta}\left[ \mat J_{\rm F}(\vec x,\vec\theta) \right] + \mathbb E_{z|\theta}\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec z|\vec\theta) \right]
\end{align}
where $\mat J_{\rm F}$ is the classical Fisher information matrix for deterministic parameters. Although we may place prior statistics on the unknown parameters, $\vec\theta$, for estimation purposes, that framework would not allow us to analyse the estimation uncertainty of particular parameter combinations.

\subsection{Summary}

\begin{table}[H]
\caption{Different bound formulations}
\label{tbl:bound_types}
\centering
\begin{tabular}{p{0.2\textwidth}p{0.4\textwidth}p{0.4\textwidth}}
  \toprule
  Bound & Details & Output  \\
  \midrule
Bayesian CRB & Features \begin{itemize}
     \item Prior on $z$ 
	\item Prior on $\theta$
	\item Singular Q
  \end{itemize} & Bound on $z$ and $\theta$ for different $R$, Number of samples, models, and uncertainty in unknown parameters (some notion of robustness to parameter estimates?)  \\
Hybrid CRB & Features \begin{itemize}
     \item Prior on $z$ 
	\item Deterministic $\theta$
	\item Singular Q
  \end{itemize} & Bound on $z$ and $\theta$ for specific $\theta$  \\
  \bottomrule
\end{tabular}
\end{table}

\section{Hybrid example}

\todo Need to derive recursion for hybrid state vector

\section{Simple hybrid example}

Consider the simple model $y_i = z t_i + \theta + w_i$ where $w_i$ is noise
\begin{align}
	\vec y = \left[\begin{array}{cc}
	 				 1 & t_1 \\ \vdots & \vdots \\ 1 & t_n 
					\end{array}\right] \left[\begin{array}{c}
					 				 \theta \\ z
									\end{array}\right] 	+ \vec w
\end{align}
Here $z$ is random parameter with prior $p_0 = \mathcal N(z_0,P_0)$, $\theta$ is deterministic unknown parameter, $w_i$ is random noise with distribution $\mathcal N(0,R)$.

The information matrix is
\begin{align}
	J = J_{\rm D} + J_{\rm P}
\end{align}
where
\begin{align}
	J_{\rm P} = \left[\begin{array}{cc}
	 				 0 & 0 \\ 0 & P_0^{-1} 
					\end{array}\right]
\end{align}
The likelihood is
\begin{align}
	\log p(\vec y|\vec x) &= c_1 -\frac{1}{2}(\vec y - \mat H\vec x)^T \mat R^{-1} (\vec y - \mat H\vec x) \\
-\nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y|\vec x) &= \mat H^T\mat R^{-1} \mat H \\
J_{\rm D} &= \mat H^T\mat R^{-1}\mat H
\end{align}

@@

Let $\vec x_k = [a_k\quad b_k\quad g]^T$ where $a_k$ is the pendulum angle and $b_k$ is the angular velocity and $g$ is an unknown deterministic parameter.
\begin{align}
	\vec f(\vec x_{k}) &= \left[\begin{array}{c}
								a_k + b_k\Delta \\ b_k - g\sin(a_k)\Delta \\ g
								\end{array}\right]
\end{align}
The linearisation used for the EKF is
\begin{align}
	\mat F(\vec x_{k}) &= \left[\begin{array}{ccc}
					1 & \Delta & 0\\
					-g\cos(a_k)\Delta & 1 & -\sin(a_k)\Delta\\
					0 & 0 & 1
				\end{array}\right]
\end{align}
The observation matrix is $H=[1\quad 0\quad 0]$, so that only the first state element is measured.
